Tools for reproducible ex-post survey data harmonization

### 2020-07-16 data-raw/article.Rmd
Documenting the recent developments

The haven package has already adopted the importing of user-defined missing values, or the user defined missing value range with the creation of the class [haven_labelled_spss](https://haven.tidyverse.org/reference/labelled_spss.html).  While this class generally works well with the older [haven_labelled](https://haven.tidyverse.org/reference/labelled.html) methods, it does not have many implemented methods in haven.  In our case, we had to make sure that when we concatenate two vectors with user-defined missing values or ranges, they will remain consistent after the join. Our class relies on the [vctrs](https://vctrs.r-lib.org/) package, which contains helper methods for creating new vector classes [cf. @R-vctrs], and which serves as the foundation of haven's own class system, too.

Our choice in this case is the that we apply integer values, long out of the range of categorical integer values in the range `99900:99999` for a consistent numeric coding of the values to be treated as missing, and we label them consistently.  The original missing labels and missing values as preserved as an attribute in our [labelled_spss_survey](http://retroharmonize.satellitereport.com/reference/labelled_spss_survey.html) class.

The consistent, harmonized treatment of the missing values allows us to easily implement the [as_numeric, as_factor and as_character](http://retroharmonize.satellitereport.com/reference/labelled_spss_survey.html) methods, by replacing the user-defined missing values with type-specific `NA` values in the R language (`NA_real_` or `NA_character_`). After this conversion, R artithmetic and statistical functions can be applied without compromising reproducability --- all metadata is present for a consistent export to SPSS or STATA, or to document, or even revert to an earlier labelling.

@@

>>>>

The haven package has already adopted the importing of user-defined missing values, or the user defined missing value range with the creation of the class [haven_labelled_spss](https://haven.tidyverse.org/reference/labelled_spss.html).  While this class generally works well with the simpler [haven_labelled](https://haven.tidyverse.org/reference/labelled.html) methods, it does not have many implemented methods in haven.  In our case, we had to make sure that when we concatenate two vectors with user-defined missing values or ranges, they will remain consistent after the join. Our [retroharmonize_labelled_spss_survey](http://retroharmonize.satellitereport.com/articles/labelled_spss_survey.html) class is an extension of the [haven_labelled_spss](https://haven.tidyverse.org/reference/labelled_spss.html). It relies on the [vctrs](https://vctrs.r-lib.org/) package, which contains helper methods for creating new vector classes [cf. @R-vctrs], and which serves as the foundation of haven's own class system, too.

Our choice in this case is the that we apply integer values, long out of the range of categorical integer values in the range `99900:99999` for a consistent numeric coding of the values to be treated as missing, and then we label them consistently.  The original missing labels and missing values as preserved as an attribute in our [labelled_spss_survey](http://retroharmonize.satellitereport.com/reference/labelled_spss_survey.html) class. We add missing labels to data that does not even have missing values for consistency.

The [haven_labelled_spss](https://haven.tidyverse.org/reference/labelled_spss.html) does not have many implemented coercion and casting methods, because there is no objective way of handling the different missing values. We had to make some subjective and consistent choices to allow a consistent re-casting with [as_numeric, as_factor and as_character](http://retroharmonize.satellitereport.com/reference/labelled_spss_survey.html) methods. For full reproducability, before replacing the user-defined missing values with the type-specific `NA` values in the R language (`NA_real_` or `NA_character_`), we carefully record the labels of these conversions, and even their pre-harmonization, historical labelling, too. After this conversion, R artithmetic and statistical functions can be applied without compromising reproducability --- all metadata is present for a consistent export to SPSS or STATA, or to document, or even revert to an earlier labelling.

The lack of clear coercion and casting disallows the concatenation or binding of [haven_labelled_spss](https://haven.tidyverse.org/reference/labelled_spss.html) vectors in a data-frame. Our subjective choice of pushing all user-defined missing values to the numeric range of range `99900:99999` with the same labelling, and applying exactly the same coding and labelling of the valid range allows us to use [vctrs::vec_c](https://vctrs.r-lib.org/) and [vctrs::vec_rbind](https://vctrs.r-lib.org/) for binding the variables together.

@@

The [document_survey_item](http://retroharmonize.satellitereport.com/reference/document_survey_item.html) function exports the entire history of the harmonization process along a codebook for the variable. ...  the looping version is not yet desinged....
